---
title: "GSS challange More Data"
author: "Sarah Medina"
date: "10/04/2022"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```
<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods
 The types of variables in this problem consist of Race and capital punishment.
 These are both categorical variables. Therefore, it is appropriate to use the 
 Cat Cat helper file. Race consists of three factors and Capital punishment
 consists of two factors. 
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~race + cappun,data=mydata)
barchartGC(~race + cappun,data=mydata, type="percent")

```


The plot demonstrates that attitudes towards capital punishment are dependent on race. The plot shows that about twice as many white people are in favor of capital punishment. Additionally, the plot shows that about 1/3 more of African Americans are opposed to Capital punishment. Lastly, for all other races the plot shows that about 1/5 more of the people included in this category are in favor of capital punishment.

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```

The numerical results demonstrate that 65 percent of white people favor capital punishment. Furthermore, the numerical results show that 58 percent of black people oppose capital punishment. Lastly, for all other races 55 percent are in favor of capital punishment. It seems that white people are more likely to be in favor of capital punishment followed by other races, and lastly black people White people are the most in favor of Capital punishment unlike the black people who are more against capital punishment.  
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->




## Now with More data



## Inferential Results


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P value is 0. This indicates that there is a statistical difference when taking into account people attitude towards capital punishment and their race.
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

## More data version
`
```{r}
fisherXY(mydata$race,mydata$cappun)
```
The fishers xy function shows there is a statistical difference on the views towards
capital punishment between the White and Black people. When we got more data the  p value of each race category indicates that there is a statistical difference between each race. The biggest difference is with the White people and the black because they have the lowest p value.



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
