---
title: "GSS challange Less Data"
author: "Sarah Medina"
date: "10/04/22"
output: html_document
---







```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,sex,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum)
mydata <- na.omit(mydata)
source("fisherXY.R")

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

In this variable analysis we will be using graphical, numerical, and inferential results to determine if race and attitudes towards capital punishment are dependent or independent form on another. The variables involved are categorical explained by categorical so the appropriate helper file will be Cat Cat. 
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~race + cappun,data=mydata)
barchartGC(~race + cappun,data=mydata, type="percent")

```
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

The bar chart shows that a little over 400 of the white people are in favor of capital punishment and that about 200 of the white people are in opposition of capital punishment. Furthermore, for the black people the bar chart shows that about the same number are in favor of capital punishment as those who are opposed to it. About 50 of the black people are in favor and about 60 are opposed. Lastly, for other races the bar chart demonstrates that about 50 are in favor of capital punishment and about 30 are against capital punishment.



## Now with LESS data (more variables give less data due to NA's) create the approate graphic displays to answer the question.





### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that 68 percent of white people are in favor of capital punishment. 32 percent of white people are opposed to capital punishment. 47 percent of black people are in favor of capital punishment and 53 percent off black people are opposed to capital punishment.Lastly, for other races 61 percent are in favor and 39 percent are against capital punishment.


## Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

```{r}
fisherXY(mydata$race,mydata$cappun, printAll = TRUE)
```

When we ran the fisher xy test with less data it indicated that the only statistical difference was between the Black and White people. The fisher xy results with less data showed that the only categories with a P value less than .05 were between the Back people and White people. The p value for white and other races was 0.2565 which is not low enough to reject the null hypothesis. Furthermore, the P value between Black and other people is 0.06299 which is also not low enough to reject the null hypothesis. 



<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`



# Discussion and Conclusion

When we ran the fisher xy test in the second branch that had less data than the master branch the p values found only a difference between the Black and White people. This demonstrates that more data is better because the master branch that had more data found a statistically difference between all the races. With less data only a statistical difference between the black and White people appears.

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
